---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "man/figures/README-",
  fig.retina = 2
)
```

# tbrf

[![Travis build status](https://travis-ci.org/mps9506/tbrf.svg?branch=master)](https://travis-ci.org/mps9506/tbrf)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/mps9506/tbrf?branch=master&svg=true)](https://ci.appveyor.com/project/mps9506/tbrf)
[![Coverage status](https://codecov.io/gh/mps9506/tbrf/branch/master/graph/badge.svg)](https://codecov.io/github/mps9506/tbrf?branch=master)
 
[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
 
The goal of tbrf is to provide time-window based rolling statistical functions. The package differs from other rolling statistic packages because the intended use is for irregular measured data. Althogh tbrf can be used to apply statistical functions to regularly sampled data, [`zoo`](https://CRAN.R-project.org/package=zoo), [`RcppRoll`](https://cran.r-project.org/package=RcppRoll), and other packages provide fast, efficient, and rich implementations of rolling/windowed functions.

An appropriate example case is water quality data that is measured at irregular time intervals. Regulatory compliance is often based on a statistical average measure or exceedance probability applied to all samples collected in the previous 7-years. tbrf can be used to display regulatory status at any sample point.

tbrf identifies the previous n measurements within the specified time window, applies the function, and outputs a varaible with the result of the rolling statistical measure.

## Installation

tbrf is still under active development (do not expect stable behavior) and can be installed from github with:

``` r
devtools::install.github("mps9506/tbrf")
```

## Available Functions

- `tbr_binom`: Rolling binomal probability with confidence intervals.

- `tbr_gmean`: Rolling geometric mean with confidence intervals.

- `tbr_mean`: Rolling mean with confidence intervals.

- `tbr_median`: Rolling median with confidence intervals.

- `tbr_misc`: Accepts user specified function.

- `tbr_sd`: Rolling standard deviation.

- `tbr_sum`: Rolling sum.

## Examples

Visualize the number of samples included in a tbrf time window:

```{r tbr_misc, fig.retina=2, message=FALSE, warning=FALSE}
library(tbrf)
library(dplyr)
library(ggplot2)

# Some sample data
df <- data_frame(date = sample(seq(as.Date('2000-01-01'),
                                   as.Date('2005-12-30'), by = "day"), 25)) %>%
  bind_rows(data.frame(date = sample(seq(as.Date('2009-01-01'),
                                         as.Date('2011-12-30'), by = "day"), 25))) %>%
  arrange(date) %>%
  mutate(value = 1:50)

# Use length function in tbr_misc to calculate n samples used in the rolling function
df <- df %>%
  tbr_misc(x = value, tcolumn = date, unit = "years", n = 5, func = length)

ggplot(df) +
  geom_point(aes(date, value)) +
  geom_errorbarh(aes(xmin = min_date, xmax = max_date, 
                     y = value, color = results)) +
  scale_color_distiller(type = "seq", palette = "OrRd", 
                        direction = 1) +
  guides(color = guide_colorbar(title = "Number of samples")) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = "Sample Date", y = "Sample Value",
       title = "Window length and n used by tbrf",
       caption = "Lines depict time window used in the function\nColors indicate number of samples in the time window")
  
```


A more general use case, visualize the 5 year geometric mean:

```{r dissolved_oxygen, fig.retina=2, message=FALSE, warning=FALSE}
data("Dissolved_Oxygen")

df <- Dissolved_Oxygen %>%
  tbr_mean(x = Average_DO, tcolumn = Date, unit = "years", n = 5)

ggplot(df) +
  geom_point(aes(Date, Average_DO)) +
  geom_line(aes(Date, mean)) +
  geom_ribbon(aes(Date, ymin = lwr.ci, ymax = upr.ci), alpha = 0.5) +
  theme_minimal() +
   labs(x = "Sample Date", y = "Concentration (mg/L)",
       title = "5-yr rolling mean",
       caption = "Line depicts the 5-yr mean\nShaded area indicates the 95% CI")
```

## Speed

The statistical functions for calculating binomial probabilities, geometric mean, mean, and median include confidence intervals by default. All except the binomial probability are currently calculated using bootstrap methodology by default. You will see substantial decrease in computing times with the following argument: `method = "classic"`. Or by using base functions with `tbr_misc`.

```{r}
data("Dissolved_Oxygen")

system.time(df <- Dissolved_Oxygen %>%
              tbr_mean(x = Average_DO, tcolumn = Date, unit = "years", n = 5, method = "boot"))

system.time(df <- Dissolved_Oxygen %>%
  tbr_mean(x = Average_DO, tcolumn = Date, unit = "years", n = 5, method = "classic"))

system.time(df <- Dissolved_Oxygen %>%
  tbr_misc(x = Average_DO, tcolumn = Date, unit = "years", n = 5, func = mean))
```


## Contributing

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By participating in this project you agree to abide by its terms.

## License

tbrf code is released under GPL-3 | LICENSE.md

`binom_ci()` is an implementation of code licensed under GPL (>=2) by Frank Harrell's [`Hmisc`](https://github.com/harrelfe/Hmisc) package.

`gmean_ci()`, `mean_ci()`, `median_ci()`,`SignTest()`, and `Winsorize()` are implementations of code licensed under GPL (>= 2) by Andri Signorell's [DescTools](https://cran.r-project.org/package=DescTools) package.

## Test Results

```{r message=FALSE, warning=FALSE, error=FALSE}
library(tbrf)

date()

devtools::test()
```

