---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "man/figures/README-",
  fig.retina = 2
)
```
# tbrf

The goal of tbrf is to provide time-window based rolling statistical functions. The package differs from other rolling statistic packages because the intended use is for irregular measured data.

An example use case is water quality data that is measured at irregular time intervals. Regulatory compliance is often based on a statistical average measure or exceedance probability applied to all samples collected in the previous 7-years.

tbrf identifies the previous n measurements within the specified time window, applies the function, and outputs a varaible with the result of the rolling statistical measure.

## Installation

tbrf is still under active development (do not expect stable behavior) and can be installed from github with:

``` r
devtools::install.github("mps9506/tbrf")
```

## Examples

tbrf includes built-in functions to calculate binomial probability (`tbr_binom`), geometric mean (`tbr_gmean`), mean (`tbr_mean`), median (`tbr_median`), standard deviation (`tbr_sd`), and sum (`tbr_sum`).

A generic function (`tbr_misc`) is provided for users to apply other functions that accept a vector of numeric values as an argument.

```{r tbr_misc, fig.retina=2, message=FALSE, warning=FALSE}
library(tbrf)
library(dplyr)
library(ggplot2)

# Some sample data
df <- data_frame(date = sample(seq(as.Date('2000-01-01'),
                                   as.Date('2005-12-30'), by = "day"), 25)) %>%
  bind_rows(data.frame(date = sample(seq(as.Date('2009-01-01'),
                                         as.Date('2011-12-30'), by = "day"), 25))) %>%
  arrange(date) %>%
  mutate(value = 1:50)

# Use length function in tbr_misc to calculate n samples used in the rolling function
df <- df %>%
  tbr_misc(x = value, tcolumn = date, unit = "years", n = 5, func = length)
df

ggplot(df) +
  geom_point(aes(date, value)) +
  geom_errorbarh(aes(xmin = min_date, xmax = max_date, 
                     y = value, color = results)) +
  scale_color_distiller(type = "seq", palette = "OrRd", 
                        direction = 1) +
  guides(color = guide_colorbar(title = "Number of samples")) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = "Sample Date", y = "Sample Value",
       title = "Window length and n used by tbrf",
       caption = "Lines depict time window used in the function\nColors indicate number of samples in the time window")
  
```



### Rolling Geometric Mean
```{r example, fig.retina=2, echo=TRUE}
library(tbrf)
library(dplyr)
library(ggplot2)

## Generate sample data
set.seed(100)
df <- data.frame(date = sample(seq(as.Date('2000-01-01'), 
                                   as.Date('2017-12-30'), by = "day"), 68),
                 value = rexp(68, 1/100))


head(df)

## apply the geomean to each row
df <- tbr_gmean(df, x = value, tcolumn = date,
                unit = "years", n = 5, conf.level = 0.95)
df


## tidy pipeline sample. sample data with breaks
set.seed(100)

df <- data.frame(date = sample(seq(as.Date('2000-01-01'), 
                                   as.Date('2007-12-30'), by = "day"), 25),
                 value = rexp(25, 1/100)) %>%
  bind_rows(data.frame(date = sample(seq(as.Date('2009-01-01'), 
                                         as.Date('2011-12-30'), by = "day"), 5),
                       value = rexp(5, 1/1000)))

df <- df %>%
  tbr_gmean(x = value, tcolumn = date, unit = "years",
            n = 5, conf.level = 0.95)
df

ggplot(df) +
  geom_point(aes(date, value)) +
  geom_line(aes(date, mean)) +
  geom_ribbon(aes(x = date, ymin = lwr.ci, ymax = upr.ci), alpha = 0.4) +
  scale_y_log10()
```

## Contributing

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By participating in this project you agree to abide by its terms.


### Test Results

```{r message=FALSE, warning=FALSE, error=FALSE}
library(tbrf)

date()

devtools::test()
```
