---
title: "tbrf Introduction"
author: "Michael Schramm"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tbrf Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tbrf)
library(dplyr)
library(ggplot2)
set.seed(1014)
```

The tbrf package aims to provide functions that return rolling or moving statistical functions based on a user specified temporal time windows (eg. 1-year, 6-months, 5-hours, etc.). This package differs from most time-series analysis packages in R that rely on applying functions to a specific number of observations.

## Introduction

Currently tbrf provides functions to calculate binomial probability, geometric mean, mean, median, standard deviation, and sum. There is also a function to apply abritary R functions. This vignette demonstrates how time-windows are applied to irregularly spaced data and each of the functions.

## Basic usage

tbrf requires an input dataframe with two variables. First, a column with times or date-times formatted as class "`POSIXt`" or "`Date`". Second, a column of observed values to calculate the statistic on. The package includes a suitable sample dataset:

```{r}
data("Dissolved_Oxygen")

head(Dissolved_Oxygen)
```

Core functions include five arguments.
```
.tbl = dataframe used by the function
x = column containing the values to calculate the statistic on
tcolumn = formatted date-time or date column
unit = character indicating the time unit used, one of "years", "months", "weeks", "days", "hours", "minutes", "seconds"
n = numeric, indicating the window length
```

If we want a 10-year rolling mean for the `Dissolved_Oxygen` dataset:

```{r}
tbr_mean(Dissolved_Oxygen, x = Average_DO,
         tcolumn = Date, unit = "years", n = 10)
```

We can use a tidy workflow:
```{r}
Dissolved_Oxygen %>%
  group_by(Station_ID) %>%
  tbr_mean(Average_DO, Date, "years", 10)
```


## Time windows



Generate some sample data:

```{r}
# Some sample data
df <- data_frame(date = sample(seq(as.Date('2000-01-01'),
                                   as.Date('2005-12-30'), by = "day"), 25)) %>%
  bind_rows(data.frame(date = sample(seq(as.Date('2009-01-01'),
                                         as.Date('2011-12-30'), by = "day"), 25))) %>%
  arrange(date) %>%
  mutate(value = 1:50)
```

We can visualize the data captured in each rolling time window using `tbr_misc()` and the `base::length()`:

```{r, fig.width=7}
df %>%
  tbr_misc(x = value, tcolumn = date, unit = "years", n = 5, func = length) %>%
  ggplot() +
  geom_point(aes(date, value)) +
  geom_errorbarh(aes(xmin = min_date, xmax = max_date, 
                     y = value, color = results)) +
  scale_color_distiller(type = "seq", palette = "OrRd", 
                        direction = 1) +
  guides(color = guide_colorbar(title = "Number of samples")) +
  theme(legend.position = "bottom") +
  labs(x = "Sample Date", y = "Sample Value",
       title = "Window length and n",
       caption = "Lines depict width of samples included in the time window\nColors indicate number of samples in the time window")
```

## Examples

### Binomial Probability

Plot the binomial probability that dissolved oxygen fell below 5 mg/L during the previous 7-year period:

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data("Dissolved_Oxygen")

Dissolved_Oxygen %>%
  mutate(x = case_when(
    Average_DO >= 5 ~ 0,
    Average_DO < 5 ~ 1)) %>%
  tbr_binom(x, Date, "years", 7, alpha = 0.05) %>%
  ggplot() +
  geom_line(aes(x = Date, y = PointEst)) +
  geom_ribbon(aes(x = Date, ymin = Lower, ymax = Upper), alpha = 0.5)

```


### Geometric Mean

Plot the rolling 7-year geometric mean:

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggalt)

data_frame(date = sample(seq(as.Date('2001-01-01'),
                                   as.Date('2017-12-31'), 
                                   by = "day"), 
                               60),
                 x = rexp(60, 1/1000)) %>%
  tbr_gmean(x, date, "years", 7, conf = 0.95, type = "perc") %>%
  ggplot() +
  geom_point(aes(date, x), alpha = 0.5) +
  geom_step(aes(date, mean)) +
  geom_ribbon(aes(x = date, ymin = lwr_ci, ymax = upr_ci), alpha = 0.5, stat = "stepribbon") +
  scale_y_log10()


```

### Mean

Plot the rolling 7-year mean:
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
Dissolved_Oxygen %>%
  mutate(Station_ID = as.factor(Station_ID)) %>%
  group_by(Station_ID) %>%
  tbr_mean(Average_DO, Date, "years", 7, conf = 0.95, type = "perc") %>%
  ggplot() +
  geom_point(aes(Date, Average_DO, color = Station_ID), alpha = 0.5) +
  geom_step(aes(Date, mean, color = Station_ID)) +
  geom_ribbon(aes(x = Date, ymin = lwr_ci, ymax = upr_ci, fill = Station_ID), alpha = 0.5, stat = "stepribbon")
```

### Median

Plot the rolling 7-year median:
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
Dissolved_Oxygen %>%
  mutate(Station_ID = as.factor(Station_ID)) %>%
  group_by(Station_ID) %>%
  tbr_median(Average_DO, Date, "years", 7, conf = 0.95, type = "perc") %>%
  ggplot() +
  geom_point(aes(Date, Average_DO, color = Station_ID), alpha = 0.5) +
  geom_step(aes(Date, median, color = Station_ID)) +
  geom_ribbon(aes(x = Date, ymin = lwr_ci, ymax = upr_ci, fill = Station_ID), alpha = 0.5, stat = "stepribbon")
```

### Generic functions

`tbr_misc()` is included to apply functions that accept a single vector of values.

For example, identify the minimum values during the previous 7 year time periods:
```{r}
Dissolved_Oxygen %>%
  tbr_misc(Average_DO, Date, "years", 7, func = min) %>%
  ggplot() +
  geom_point(aes(Date, Average_DO), alpha = 0.5) +
  geom_line(aes(Date, results))
```

### Standard Deviation

Plot the rolling 7-year SD:

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
Dissolved_Oxygen %>%
  tbr_sd(Average_DO, Date, "years", 7) %>%
  ggplot() +
  geom_line(aes(Date, sd))
```

### Sum

Plot the rolling 7-year sum:

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
Dissolved_Oxygen %>%
  mutate(Station_ID = as.factor(Station_ID)) %>%
  group_by(Station_ID) %>%
  tbr_sum(Average_DO, Date, "years", 7) %>%
  ggplot() +
  geom_line(aes(Date, sum, color = Station_ID))
```


## Units

Allowable character values for `unit` include `c("years", "months", "weeks", "days", "hours", "minutes", "seconds")`. Example using `"minutes"` and `"hours"`:

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
y = 3 * sin(2 * seq(from = 0, to = 4*pi, length.out = 100)) + rnorm(100)
time = sample(seq(as.POSIXct(strptime("2017-01-01 00:01:00", "%Y-%m-%d %H:%M:%S")),
                  as.POSIXct(strptime("2017-01-01 23:00:00", "%Y-%m-%d %H:%M:%S")),
                  by = "min"), 100)

df <- data_frame(y, time)

df %>%
  tbr_mean(y, time, "minutes", n = 30) %>%
  ggplot() +
  geom_point(aes(time, y)) +
  geom_line(aes(time, mean))


df %>%
  tbr_mean(y, time, "minutes", n = 60) %>%
  ggplot() +
  geom_point(aes(time, y)) +
  geom_line(aes(time, mean))


df %>%
  tbr_mean(y, time, "hours", n = 5) %>%
  ggplot() +
  geom_point(aes(time, y)) +
  geom_line(aes(time, mean))

```

## CI method

Confidence intervals in `tbr_gmean`, `tbr_mean`, and `tbr_median` are calculated using `boot_ci`. If you do not need confidence intervals, calulcation times are substantially shorter. `parallel`, `ncores`, and `cl` arguments are passed to `boot` and can improve computation times. An example using parallel processing for Windows systems is below:


```{r eval=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(snow)

cl <- makeCluster(4, type = "SOCK")

tbr_mean(Dissolved_Oxygen, Average_DO, Date, 
         "years", 5, R = 1000, conf = .95,
         type = "perc", parallel = "snow", 
         cl = cl)

stopCluster(cl)
```
